---
title: "Calculate_AUC"
author: "Xuhong"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We set the initial values and randomly generated x and y,We also calculated the numbers of y=1 and y=0,
The number of cases is 724, indicating y=1.
The number of controls is 2276, indicating y=0.
```{r 1}
beta0 <- -4
beta1 <- 0.6
# generate x  y_true
set.seed(0)
x <- rnorm(3000, mean = 10, sd = 4)
p <- exp(beta0 + beta1 * x) / (1 + exp(beta0 + beta1 * x))

# get the true label from two-points Bernoulli
set.seed(1)
y_true <- rbinom(length(x), 1, p)
head(y_true, 100)

count_y_equals_1 <- sum(y_true == 1)
count_y_equals_0 <- sum(y_true == 0)
```

Subsequently, the labels y and x, as determined earlier, were employed to perform logistic regression, yielding the parameter estimates:
Beta0_hat=-3.95,
Beta1_hat=0.602.
These estimated values were then utilized to calculate the predicted values.
```{r 2}
# Fitting logistic model
data <- data.frame(x, y_true)
log_model <- glm(y_true ~ x, data = data, family = binomial(link="logit"))
summary(log_model)

#get the estimate value of intercept and slope
coefficients <- coef(log_model)
beta0_hat <- coefficients["(Intercept)"]
beta1_hat <- coefficients["x"]

#get the predicate value logistic model 
y_pred<-exp(beta0_hat + beta1_hat * x) / (1 + exp(beta0_hat + beta1_hat * x))
```
  
Moving forward, we proceed to compute the TPR and FPR. Subsequently, we construct the ROC curve. It is noteworthy that we systematically vary the threshold from 0 to 1 in increments of 0.01 during this process.
```{r 3}
#ROC plot method 1 and print the value of AUC
library(pROC)
moni<-roc(y_true~y_pred,smooth=FALSE)
plot(moni, print.auc=TRUE, auc.polygon=TRUE, 
     grid=c(0.1,0.1),
     grid.col=c("#BAC94A",'#329D9C'),
     max.auc.polygon=TRUE,
     auc.polygon.col="#CEE4D2")

#ROC curve plot method 2
cutoff <- seq(0, 1, length = 100)
FPR <- sapply(cutoff,
              function(thresh) {
                sum(y_pred >= thresh & y_true != 1) / sum(y_true != 1)
              })
TPR <- sapply(cutoff,
              function(thresh) {
                sum(y_pred >= thresh & y_true == 1) / sum(y_true == 1)
              })
```

```{r 4, include=FALSE}
# plot ROC curve
plot(FPR, TPR, type = "l",lwd = 2, xlab = "False Positive Rate", ylab = "True Positive Rate", main = "ROC Curve")
lines(FPR, TPR, col = "#C5D9B3", lwd = 2)

grid()
abline(h = seq(0, 1, by = 0.1), v = seq(0, 1, by = 0.2), col = "#D6C1A9", lty = 2)
```

In this part, we present three methods for calculating the AUC, building upon the aforementioned procedures.

```{r 5}
#method 1
library(pROC)
auc_value<-auc(y_true,y_pred)
```
The value of AUC based on method 1 is 0.897

```{r 6}
#method 2
s <- 0
for (i in which(y_true == 1)) {
  for (j in which(y_true == 0)) {
    if (y_pred[i] > y_pred[j]) {
      s <- s + 1
    } 
  }
}
s <- s / (sum(y_true == 1) * sum(y_true == 0))
s
#consider the case of equality
s1 <- 0
equal_count <- 0
for (i in which(y_true == 1)) {
  for (j in which(y_true == 0)) {
    if (y_pred[i] > y_pred[j]) {
      s1 <- s1 + 1
    } else if (y_pred[i] == y_pred[j]){
      s1<-s1+0.5
      equal_count <- equal_count + 1 
    }
  }
}
s1 <- s1 / (sum(y_true == 1) * sum(y_true == 0))
s1
```
The value of AUC based on method 2 is 0.89653931


```{r 7}
#method 3
#number lattice
library(ggplot2)
# sorted y_pred，get the indices
sorted_indices <- order(y_pred, decreasing = TRUE)
sorted_y_pred <- y_pred[sorted_indices]
sorted_y_true <- y_true[sorted_indices]

# innitialize x and y as 0
x <- 0
y <- 0

# creat a data frame
df <- data.frame(x = numeric(length(sorted_y_pred)), y = numeric(length(sorted_y_pred)))

# traverse the sorted y_pred
for (i in 1:length(sorted_y_pred)) {
 
  if (sorted_y_true[i] == 1) {
    y <- y + 1
  } else {
    x <- x + 1
  }
  
  # update data frame
  df[i, ] <- c(x, y)
}

# use ggplot2 
ggplot(df, aes(x, y)) +
  geom_point(color='#A5CAD2',size=0.1) +
  labs(x = "FPR", y = "TPR", title = "ROC curve")+
  theme_minimal()+ theme(panel.grid.major = element_line(color = "#ECD59F", linetype = "dashed"))


# get the coordinate of x and y
x_values <- df$x
y_values <- df$y

# calculate the area under curve
area_under_curve <- integrate(Vectorize(function(x) approx(x_values, y_values, xout = x)$y), min(x_values), max(x_values))$value

# calculate the total 
total_area <- max(x_values) * max(y_values)

# calculate the auc
auc_area <- (area_under_curve / total_area)
```

